---
title: "A (Not Entirely Novel) Problem with No Definitive Solutions"
description: |
  How can we extract meaningful information from a corpora of Renaissance texts using topic modeling?
author:
  - name: Cassidy All
    affiliation: University of Colorado Boulder
    affiliation_url: https://example.com/spacelysprokets
date: 09-27-2023
bibliography: sample.bib
csl: ieee.csl
output:
  distill::distill_article:
    toc: true
    self_contained: false
---

# Introduction

Digital humanities work promises researchers access to unthinkably large corpora spanning upwards of 278,000 books @gao2020 and 300 billion words @goldberg2013. Computer scientists have effectively used these corpora to train increasingly-powerful Artificial Intelligence (AI) models. However, many of the insights in these corpora remain inscrutable to human researchers.

Blei @blei2012 and others have described some of the valuable research which can arise from corpora-level analysis centered in the humanities. Corpora-level analysis allows researchers to examine the evolution of texts over time; search for specific topics; understand the influence of authorship, provenance, and other non-contextual features on literature's themes; and more.

A major tool in doing this digital humanities work is topic modeling^[A probabilistic approach to generating groups of words and phrases from a collection of texts.], however, its difficulty represents a major stumbling block for new and experienced scholars alike. Broadly, we consider the motivations of topic modeling, adapted from @batmanghelich2018, to be:

  * **Organizing** documents into thematic categories
  * **Designing** categories with appropriate levels of granularity and abstractness
  * **Describing** the evolution of those categories over time
  * **Enabling** a domain expert to analyze and understand the content
  * **Finding** relationships between categories
  * **Understanding** how authorship, origin, and other contextual features influence content

## Metaphors, Irony, and Context

Corpora of historical literature share much of the same problem space as more modern corpora, however, they also present unique challenges.

## Our Corpus

To explore race and politics in Renaissance literature&mdash;as well as develop new methods in the digital humanities&mdash;we are utilizing the [Visualizing English Print](https://graphics.cs.wisc.edu/WP/vep/vep-tcp-collection/) corpus with standardized spellings generated from texts released by the [Text Creation Partnership](https://textcreationpartnership.org/) as part of the [Early English Books Online](https://quod.lib.umich.edu/e/eebogroup/) corpus.

It contains ____ from ___ to ___ and ___ unique words. Explore some of the collection's metadata below.

```{r, layout="l-body-outset", echo=FALSE}
library(rmarkdown)
library(readr)
vep_metadata <- read_csv('EEBOTCP1.csv',show_col_types = FALSE)
vep_metadata <- vep_metadata[,c("Title","Date","Author","Pages","Terms")]
paged_table(vep_metadata,options = list(rows.print = 5, max.print = 30000))
```

## “Wishlist” for topic models

* Choosing an optimal number (K) of topics
* Coherent, distinct, and, non-ambiguous topics
* Reasonable compute time for mid-size(?) corpora
* Contextual understandings of KWIC

## Better Evaluation

What's going on?


# Research Directions

## Specific Work

* Produce ground truth corpus
* Understand race, politics over time

## Broad Contributions
* Conduct a comprehensive review of major research directions in topic modeling as they apply to the analysis of historical corpora
* Design and evaluate ensemble models
* Explore and design evaluation metrics from multistakeholder perspectives
* Improve the computational efficiency of the Top2Vec library
* Contribute to TopModel, OCTIS, and other topic modeling frameworks
* Develop new models and a set of best practices for understanding metaphor, irony, and other context-dependent phrases

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = FALSE)

# Learn more about creating blogs with Distill at:
# https://rstudio.github.io/distill/blog.html

```



